---
title: "ISSS603_Cust_A1"
author: "Matthias Oh"
date: "4 February 2019"
output: html_document
---

##Default initialisation
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load libraries
```{r, echo=FALSE}
packages <- c('data.table', 'scales', 'lubridate', 'GGally', 'tidyverse', 'rfm')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

##Load Dataset
```{r, echo=FALSE}
setwd("D:/Customer Analytics/A1/data")

#Set blanks, "NA", and "NULL" values to NA.
data <- fread("Assignment1 dataset Jan2019.csv", sep=",", na.strings=c("","NA","NULL"))

str(data)
```

##Rectify Erroneous Datatypes and Recode Inconsistent Formatting.
```{r, echo=FALSE}
#Change from character to date data type.
data$`Effective To Date` <- as.Date(parse_date_time(data$`Effective To Date`,'mdy'))
```

##Summary Statistics.
```{r, echo=FALSE}
summary(data)
```

##Check for missing values.
```{r}
colSums(is.na(data))
```

##Check for duplicate rows.
```{r}
data[duplicated(data)]
```

Check if there are multiple rows corresponding to the same Customer ID.
Different results will result in different methods of calculating each customer's revenue contribution to the company.
```{r}
data[duplicated(data$Customer)]
```

##NA Cleaning Steps
- Normalisation / Recoding.
- Transformation.
- Standardisation.

##Feature Engineering.
```{r}
data <- data %>%
  mutate(Revenue = `Monthly Premium Auto` * `Months Since Policy Inception`)

head(data[,c("Monthly Premium Auto", "Months Since Policy Inception", "Revenue")])
```

##Univariate Plots - Numerical.
```{r}
nums <- unlist(lapply(data, is.numeric))

data[, nums] %>%
  gather() %>%
  filter(!(key %in% c('Months Since Last Claim','Months Since Policy Inception',
                      'Number of Open Complaints','Number of Policies'))) %>%
  ggplot(aes(value)) +
  geom_histogram() +
  facet_wrap(~key, scales='free')
```

```{r}
ints <- c('Months Since Last Claim','Months Since Policy Inception',
          'Number of Open Complaints','Number of Policies')

data[, ints] %>%
  gather() %>%
  ggplot(aes(value)) +
  geom_bar() +
  facet_wrap(~key, scales='free') +
  scale_x_continuous(breaks = scales::pretty_breaks())
```

**Observations:**  
1. Based on skew, there seems to be outliers for 'Customer Lifetime Value', 'Monthly Premium Auto', and 'Total Claim Amount'.
2. There seems to be a large number of customers with 0 income, and they are likely to be unemployed.
```{r}
data %>%
  select(Income, EmploymentStatus) %>%
  filter(Income == 0) %>%
  group_by(EmploymentStatus) %>%
  summarise(subtotal = n())
```

##Check for Outliers - Numeric.
```{r}
data[, nums] %>%
  lapply(sd)
```

##Univariate Plots - Character.
```{r, fig.height=8, fig.width=10}
chars <- unlist(lapply(data, is.character))

data[, chars] %>%
  gather() %>%
  filter(!(key %in% c('Customer','Effective To Date'))) %>%
  ggplot(aes(value)) +
  geom_bar() +
  facet_wrap(~key, scales='free') + 
  theme(axis.text.x=element_text(angle=90))
```

**Observations:**
1.Looking at historical 'Response', the sample dataset is imbalanced.

##Univariate Plots - Date.
```{r}
data %>%
  ggplot(aes(`Effective To Date`)) +
  geom_freqpoly() +
  scale_x_date(labels=date_format("%d/%m/%y"),breaks = "7 day") + 
  theme(axis.text.x=element_text(angle=90))
```

##Multivariate Plots.
CLV -vs- Revenue
```{r, echo=FALSE}
data %>%
  ggplot(aes(x=Revenue, y=`Customer Lifetime Value`)) +
  geom_point() +
  geom_smooth()
```


```{r, echo=FALSE}
data %>%
  ggplot(aes(x=`Policy Type`, fill=Response)) +
  geom_bar()
```

```{r, echo=FALSE}
data %>%
  group_by(`Policy Type`, Response) %>%
  summarise(Subtotal = n()) %>%
  mutate(Proportion = Subtotal/sum(Subtotal))
```

```{r, echo=FALSE}
data %>%
  ggplot(aes(x=`Sales Channel`, fill=Response)) +
  geom_bar()
```

```{r, echo=FALSE}
data %>%
  group_by(`Sales Channel`, Response) %>%
  summarise(Subtotal = n()) %>%
  mutate(Proportion = Subtotal/sum(Subtotal))
```

```{r, echo=FALSE}
data %>%
  ggplot(aes(x=`Sales Channel`, y=`Monthly Premium Auto`)) +
  geom_boxplot()
```
```{r, echo=FALSE}
data %>%
  group_by(`Sales Channel`) %>%
  summarise(MeanPrem=mean(`Monthly Premium Auto`),
            MedPrem=median(`Monthly Premium Auto`))
```

```{r, echo=FALSE}
data %>%
  ggplot(aes(x=`Policy Type`, y=`Monthly Premium Auto`)) +
  geom_boxplot()
```
```{r, echo=FALSE}
data %>%
  group_by(`Policy Type`) %>%
  summarise(MeanPrem=mean(`Monthly Premium Auto`),
            MedPrem=median(`Monthly Premium Auto`))
```

##RFM
* R: 'Months Since Policy Inception'
* F: 'Number of Policies'
* M: 'Monthly Premium Auto'
```{r}
rfm_model <- rfm_table_customer(data, customer_id = Customer, 
                                 n_transactions = `Number of Policies`,
                                 recency_days = `Months Since Policy Inception`,
                                 total_revenue = `Monthly Premium Auto`,
                                 recency_bins = 5, frequency_bins = 5,
                                 monetary_bins = 5)
rfm_result <- rfm_model$rfm
head(rfm_result)
```

```{r, echo=FALSE}
summary(rfm_result)
```

##RFM 3D Heatmap.
```{r, echo=FALSE}
rfm_heatmap(rfm_model)
```

**Observations:**
1. No customers have a 'Frequency' score of 3.
2. If <Revenue> is used as 'Monetary' of RFM, there will be a correlation between low Recency and high Monetary scores, because there will be greater premium amounts accrued over longer durations.
3. After changing 'Monetary' to use <Monthly Premium Auto> instead, which is not a function of <Months Since Policy Inception> (thereby eliminating correlation between R and M of RFM), it was noted that all segments are similar in Monetary terms.

##RFM Bar Chart.
```{r, echo=FALSE}
rfm_bar_chart(rfm_model)
```

##Compute Response Rate and Expected Revenue.
```{r, echo=FALSE}
temp <- rfm_result %>%
  full_join(select(data,Customer, Response), 
            by=c("customer_id"="Customer")) %>%
  group_by(rfm_score, Response) %>%
  summarise(NumCust = n(), Prem = sum(amount)) %>%
  mutate(RespRate = NumCust/sum(NumCust))

temp$rfm_score <- as.character(temp$rfm_score)

head(temp)
```

```{r, echo=FALSE}
temp2 <- temp %>%
  filter(Response == "Yes") %>%
  mutate(ERev = RespRate * Prem) %>%
  arrange(desc(ERev))

temp2
```

```{r, echo=FALSE}
temp2 %>%
  ggplot(aes(x=reorder(rfm_score,-ERev),y=RespRate, fill=ERev)) +
  geom_bar(stat="identity") + 
  theme(axis.text.x=element_text(angle=90)) +
  scale_fill_continuous(low="blue",high="red") +
  labs(x="RFM Score Segments",y="Response Rate",fill="Expected Revenue")
```

```{r, echo=FALSE}
temp3 <- temp %>%
  group_by(rfm_score) %>%
  summarise(SegNumCust = sum(NumCust)) 

temp2 <- temp2 %>%
  inner_join(select(temp3,rfm_score,SegNumCust),by='rfm_score')

head(temp2)
```

```{r, echo=FALSE}
indexref <- length(which(cumsum(temp2$SegNumCust) < 3000)) + 1
targetsegments <- temp2[1:indexref,]

targetsegments
```

```{r, echo=FALSE}
mean(temp2$RespRate)
mean(targetsegments$RespRate)
```

```{r, echo=FALSE}
write.csv(targetsegments,'targetsegments.csv')
```

